---
title: "Covid91 vaccine study Final2022"
author: "Michelle Benavides"
date: "August 1, 2022"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## R Code Details

First, the variables from FinalData needed to be changed from character variables into factor variables. the appropriate dplyr code was used to do this. Next, specific subgroups had to be separated from the broader categories in order to properly studied. Gal and guy were separated into subgroups, as well as gay sexual orientation and people that answered yes to participating in drug use. All of these categories are isolated to see weather or not these particular populations should or should not take the new covid-19 vaccine. In order to create the effectiveness interval for all four categories, each group of people in this study had to be stored in separate variables.

## Males

```{r}
MaleData<-filter(FinalData, sex == "M")
MaleData<-select(MaleData, -sex)
```

### Graphical Results

```{r}
dd2 <- MaleData %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

## `summarise()` has grouped output by 'infected'. You can override using the `.groups` argument.

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of males people and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.  
The frequency plot above demonstrates that almost an equal amount of the male sample took either the vaccine or the placebo. The fact that the sample had almost an equal amount of placebos or vaccine makes the study reliable and well-grounded. While the percentage plot demonstrates that about 30% of the people that took the vaccine did in fact get covid-19, while only above 60% of the people taking the placebo contracted covid-19. A substantially lower percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=MaleData)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of men that got covid-19 after taking the placebo and taking the vaccine. Among the guys, 32.22% got covid-19 after taking the vaccine, while the approximately 67.78% of guys that were given the placebo contracted covid-19. This data shows that the vaccine seems to be working significantly better than the placebo. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
MaleFisher <- fisher.test(table1)
MaleFisher
```

These results is the output of the Fisher’s Exact test. The p-value displayed above is extremely small at 2.645e-11. This is an awfully small number. seeing this number means that these results are not likely to occur by chance. The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option. The odd ratio above illustrates a ratio of 2.09. The men who received tha vaccine were two times more likely to get covid-19 than those who received the placebo.

### Effectiveness Confidence Interval

```{r}
effect <- MaleFisher$p.value/(MaleFisher$p.value +1)
MaleEffectiveness = MaleFisher$estimate/(MaleFisher$estimate +1)*100
print(MaleEffectiveness)

low=(MaleFisher$conf.int[1]/(MaleFisher$conf.int[1] +1))*100
hi=(MaleFisher$conf.int[2]/(MaleFisher$conf.int[2] +1))*100
options(digits=2)
```


Above is the effectiveness for the covid-19 vaccine for males. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting men from getting covid-19.The odds ratio is 68. The confidence interval more the male sample is 62.51 to 72.51.

## Females

```{r}
FemaleData<-filter(FinalData, sex == "F")
FemaleData<-select(FemaleData, -sex)
```

### Graphical Results

```{r}
dd2 <- FemaleData %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

## `summarise()` has grouped output by 'infected'. You can override using the `.groups` argument.

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of females and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.  
The frequency plot above demonstrates that almost an equal amount of the females in the sample took either the vaccine or the placebo. The fact that the sample had almost an equal amount of placebos or vaccine makes the study reliable and well-grounded. While the percentage plot demonstrates that about 40% of the people that took the vaccine did in fact get covid-19, while almost 60% of the people taking the placebo contracted covid-19. A substantially lower percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=FemaleData)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of women that got covid-19 after taking the placebo and taking the vaccine. Among the women, 41% got covid-19 after taking the vaccine, while the approximately 59% of women that were given the placebo contracted covid-19. This data shows that the vaccine seems to be working significantly better than the placebo. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
FemaleFisher <- fisher.test(table1)
FemaleFisher
```

These results is the output of the Fisher’s Exact test. The p-value displayed above is extremely small at 9e-04. This is an awfully small number. seeing this number means that these results are not likely to occur by chance. The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the favorable option. The odd ratio above illustrates a ratio of 1.4. The women who received the vaccine were two times more likely to get covid-19 than those who received the placebo.

### Effectiveness Confidence Interval

```{r}
effect <- FemaleFisher$p.value/(FemaleFisher$p.value +1)
FemaleEffectiveness = FemaleFisher$estimate/(FemaleFisher$estimate +1)*100
print(FemaleEffectiveness)

low=(FemaleFisher$conf.int[1]/(FemaleFisher$conf.int[1] +1))*100
hi=(FemaleFisher$conf.int[2]/(FemaleFisher$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is effective in protecting men from getting covid-19.The odds ratio is 59. The confidence interval more the male sample is 53.50 to 64.01.

## LGBTQ+

```{r}
LGBTQData<-filter(FinalData, LGBTQ == "gay")
LGBTQData<-select(LGBTQData, -LGBTQ)
```

### Graphical Results

```{r}
dd2 <- LGBTQData %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

## `summarise()` has grouped output by 'infected'. You can override using the `.groups` argument.

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of LGBTQ people and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.  
The frequency plot above demonstrates that almost an equal amount of the LGBTQ in the sample took either the vaccine or the placebo. The fact that the sample had almost an equal amount of placebos or vaccine makes the study reliable and well-grounded. While the percentage plot demonstrates that about 80% of the people that took the vaccine did in fact get covid-19, while almost 10% of the people taking the placebo contracted covid-19. A substantially lower percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of LGBTQ people that got covid-19 after taking the placebo and taking the vaccine. Among the LGBTQ people, 87% got covid-19 after taking the vaccine, while only 13% of LGBTQ people that were given the placebo contracted covid-19. This data shows that the vaccine seems to be working significantly worse than the placebo. 

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQFisher <- fisher.test(table1)
LGBTQFisher
```

These results is the output of the Fisher’s Exact test. The p-value displayed above is extremely small at 2e-09. This is an awfully small number. seeing this number means that these results are not likely to occur by chance. The null hypothesis predicted that the deaths related to covid-19 are independent of sexual orientation. Since it is clear that sexual orientation does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sexual orientation, is the favorable option. The odd ratio above illustrates a ratio of 0.15.

### Effectiveness Confidence Interval

```{r}
effect <- LGBTQFisher$p.value/(LGBTQFisher$p.value +1)
LGBTQEffectiveness = LGBTQFisher$estimate/(LGBTQFisher$estimate +1)*100
print(LGBTQEffectiveness)

low=(LGBTQFisher$conf.int[1]/(LGBTQFisher$conf.int[1] +1))*100
hi=(LGBTQFisher$conf.int[2]/(LGBTQFisher$conf.int[2] +1))*100
options(digits=2)
```

Above is the effectiveness for the covid-19 vaccine for females. If the effectiveness is above 50% that means that the vaccine is effective. Based on the interval seen above, we can conclude that the vaccine is ineffective in protecting men from getting covid-19.The odds ratio is 13. The confidence interval more the male sample is 5.65 to 23.78.

## Druggies

```{r}
DruggiesData<-filter(FinalData, DrugUser == "yes")
DruggiesData<-select(DruggiesData, -DrugUser)
```

### Graphical Results

```{r}
dd2 <- DruggiesData %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

## `summarise()` has grouped output by 'infected'. You can override using the `.groups` argument.

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of drug users and the relationship to the contraction of covid-19 by either taking the vaccine or the placebo. The codes that create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.  
The frequency plot above demonstrates that almost an equal amount of the drug users in the sample took either the vaccine or the placebo. The fact that the sample had almost an equal amount of placebos or vaccine makes the study reliable and well-grounded. While the percentage plot demonstrates that about 85% of the people that took the vaccine did in fact get covid-19, while almost 15% of the people taking the placebo contracted covid-19. A substantially lower percentage of people who took the vaccine contracted covid-19 versus those who were only given the placebo.

### Numerical Results

### Inferential Results

## Discussion and Conclusion